Assessing the potential of Jacobian-free Newton-Krylov methods for cell-centred finite volume solid mechanics

### 2024-10-22 article.tex
Sec 4.2 updates
Add Rhie-Chow appendix (but shows 2nd order)

A fully explicit procedure can also be viewed from this perspective by selecting an approximate Jacobian which is diagonal $\bb{\tilde{D}}$, making solution of the linear system trivial:
\begin{eqnarray} \label{eq:exp}
    \bb{\tilde{D}}(\bb{u}_k) \;\delta \bb{u} = -\bb{R}(\bb{u}_k)
\end{eqnarray}

@@

			\textbf{FSI} & $1\,252$ + 156 & $4\,084$ & 758 				& $4\,490$ & 732 \\
			\emph{2-D, dynamic,}	& $5\,008$ + 624 & $28\,445$ & 777 	& $48\,440$ & 760 \\
			\emph{hyperelastic} & $20\,032$ + $2\,496$ & $154\,237$ & 729 	& & \hl{re-rnng all} \\

@@

\hl{Add this for 1-D}

>>>>

A fully explicit procedure can also be viewed from this perspective by selecting an approximate Jacobian, which is diagonal $\bb{\tilde{D}}$, making the solution of the linear system trivial:
\begin{eqnarray} \label{eq:exp}
    \bb{\tilde{D}}(\bb{u}_k) \;\delta \bb{u} = -\bb{R}(\bb{u}_k)
\end{eqnarray}

@@

			\textbf{FSI} & $1\,252$ + 156 & $3\,830$ & 748 				& $5\,360$ & 738 \\
			\emph{2-D, dynamic,}	& $5\,008$ + 624 & $32\,074$ & 834 	& $52\,372$ & 775 \\
			\emph{hyperelastic} & $20\,032$ + $2\,496$ & $170\,919$ & 785 	& $218\,797$ & 748 \\

@@

On a 1-D uniform mesh with spacing $\Delta x$ and unity areas, the Rhie-Chow term stabilisation term (Equation \ref{eq:RhieChow}) for centre cell $P$ becomes
\begin{eqnarray}
\mathcal{D}_{\text {Rhie-Chow}}
	&=&
	\sum_{f \in N_f} \alpha \bar{K}_f
	\left[
	\left|\bb{\Delta}_f\right| \frac{ \bb{u}_{N_f} - \bb{u}_P}{\left|\bb{d}_f\right|}
	- \bb{\Delta}_f \cdot \left(\bb{\nabla} \bb{u} \right)_f
	\right]
	\left|\bb{\Gamma}_{f}\right| \notag \\
	&=& 
	\alpha \bar{K} \left[ \frac{ \bb{u}_{E} - \bb{u}_P}{\Delta x} - \left(\bb{\nabla} \bb{u} \right)_e \right]
	+ \alpha \bar{K} \left[ \frac{ \bb{u}_{W} - \bb{u}_P}{\Delta x} + \left(\bb{\nabla} \bb{u} \right)_w \right]
	\notag \\
\end{eqnarray}
where $E$ and $W$ indicate the east and west neighbour cell centre values, $EE$ and $WW$ are the far east and west neighbour cell centre values, and $e$ and $w$ indicate east and west face values;
$\alpha$ and $\bar{K}$ are assumed uniform.
The face gradients are calculated as
\begin{eqnarray}
	\left(\bb{\nabla} \bb{u} \right)_e = \frac{1}{2} \left[ \left(\bb{\nabla} \bb{u} \right)_P + \left(\bb{\nabla} \bb{u} \right)_E \right]  \notag \\
	\left(\bb{\nabla} \bb{u} \right)_w = \frac{1}{2} \left[ \left(\bb{\nabla} \bb{u} \right)_W + \left(\bb{\nabla} \bb{u} \right)_P \right]
\end{eqnarray}
where the cell centre gradients are calculated as
\begin{eqnarray}
	\left(\bb{\nabla} \bb{u} \right)_W = \frac{\bb{u}_{P} - \bb{u}_{WW}}{2\Delta x} \notag \\
	\left(\bb{\nabla} \bb{u} \right)_P = \frac{\bb{u}_{E} - \bb{u}_{W}}{2\Delta x} \notag \\
	\left(\bb{\nabla} \bb{u} \right)_E = \frac{\bb{u}_{EE} - \bb{u}_{P}}{2\Delta x}
\end{eqnarray}

The final Rhie-Chow term becomes
\begin{eqnarray} \label{eq:Rhie_Chow_1D}
	\mathcal{D}_{\text {Rhie-Chow}}
	&=& 
	\alpha \bar{K}
		\left\{
		\frac{ \bb{u}_{E} - \bb{u}_P}{\Delta x}
		- \frac{1}{2} \left[ \left(\bb{\nabla} \bb{u} \right)_P + \left(\bb{\nabla} \bb{u} \right)_E \right]
	+ \frac{ \bb{u}_{W} - \bb{u}_P}{\Delta x}
		+ \frac{1}{2} \left[ \left(\bb{\nabla} \bb{u} \right)_W + \left(\bb{\nabla} \bb{u} \right)_P \right]
		\right\}
	\notag \\
	&=& 
	\alpha \bar{K}
		\left\{
		\frac{ \bb{u}_{E} - \bb{u}_P}{\Delta x}
		- \frac{1}{2} \left(\bb{\nabla} \bb{u} \right)_E
	+ \frac{ \bb{u}_{W} - \bb{u}_P}{\Delta x}
		+ \frac{1}{2} \left(\bb{\nabla} \bb{u} \right)_W 
		\right\}
	\notag \\
	&=& 
	\alpha \bar{K}
		\left[
		\frac{ \bb{u}_{E} - \bb{u}_P}{\Delta x}
%		- \frac{1}{2} \left( \frac{\bb{u}_{E} - \bb{u}_{W}}{2\Delta x} + \frac{\bb{u}_{EE} - \bb{u}_{P}}{2\Delta x} \right)
		- \frac{\bb{u}_{EE} - \bb{u}_{P}}{4\Delta x}
	+ \frac{ \bb{u}_{W} - \bb{u}_P}{\Delta x}
%		+ \frac{1}{2} \left( \frac{\bb{u}_{P} - \bb{u}_{WW}}{2\Delta x}  + \frac{\bb{u}_{E} - \bb{u}_{W}}{2\Delta x} \right)
		+ \frac{\bb{u}_{P} - \bb{u}_{WW}}{4\Delta x} 
		\right]
	\notag \\
%	&=& 
%	\alpha \bar{K}
%		\left[
%		\frac{ \bb{u}_{E} - 2\bb{u}_P + \bb{u}_W}{\Delta x}
%		- \frac{1}{2} \left(
%			\frac{\bb{u}_{E} - \bb{u}_{W}}{2\Delta x} + \frac{\bb{u}_{EE} - \bb{u}_{P}}{2\Delta x}
%			+ \frac{\bb{u}_{P} - \bb{u}_{WW}}{2\Delta x}  + \frac{\bb{u}_{E} - \bb{u}_{W}}{2\Delta x}
%		\right)
%		\right]
%	\notag \\
%	&=& 
%	\frac{\alpha \bar{K}}{4 \Delta x}
%		\left[
%		4\bb{u}_{E} - 8\bb{u}_P + 4\bb{u}_W
%		-  \bb{u}_{E} + \bb{u}_{W} + \bb{u}_{EE} - \bb{u}_{P}
%	 	+ \bb{u}_{P} - \bb{u}_{WW}  + \bb{u}_{E} - \bb{u}_{W}
%		\right]
%	\notag \\
	&=& 
	\frac{\alpha \bar{K}}{4 \Delta x}	\left[- \bb{u}_{WW} + 4\bb{u}_W - 6\bb{u}_P + 4\bb{u}_{E} - \bb{u}_{EE} \right]
\end{eqnarray}

A local truncation analysis can be performed using the following truncated Taylor series about the true solution ($\bb{U}$) and its gradients ($\left( \bb{\nabla} \bb{U} \right)_P$, $\left( \bb{\nabla}^2 \bb{U} \right)_P$, $...$) at $P$:
\begin{eqnarray} \label{eq:taylor_series_1d}
	\bb{u}_{WW} &=& \bb{U}_P - 2\Delta x \left( \bb{\nabla} \bb{U} \right)_P + \frac{4\Delta x^2}{2!} \left( \bb{\nabla}^2 \bb{U} \right)_P
		- \frac{8\Delta x^3}{3!} \left( \bb{\nabla}^3 \bb{U} \right)_P + O(\Delta x^4) \notag \\
	\bb{u}_W &=& \bb{U}_P - \Delta x \left( \bb{\nabla} \bb{U} \right)_P + \frac{\Delta x^2}{2!} \left( \bb{\nabla}^2 \bb{U} \right)_P
		- \frac{\Delta x^3}{3!} \left( \bb{\nabla}^3 \bb{U} \right)_P + O(\Delta x^4) \notag \\
	\bb{u}_P &=& \bb{U}_P \notag \\
	\bb{u}_E &=& \bb{U}_P + \Delta x \left( \bb{\nabla} \bb{U} \right)_P + \frac{\Delta x^2}{2!} \left( \bb{\nabla}^2 \bb{U} \right)_P 
		+ \frac{\Delta x^3}{3!} \left( \bb{\nabla}^3 \bb{U} \right)_P + O(\Delta x^4) \notag \\
	\bb{u}_{EE} &=& \bb{U}_P + 2\Delta x \left( \bb{\nabla} \bb{U} \right)_P + \frac{4\Delta x^2}{2!} \left( \bb{\nabla}^2 \bb{U} \right)_P 
		+ \frac{8\Delta x^3}{3!} \left( \bb{\nabla}^3 \bb{U} \right)_P + O(\Delta x^4)
\end{eqnarray}
where $O(\Delta x^4)$ are higher-order terms with a leading term porportional to $\Delta x^4$.

Substituting the expressions above (Equations \ref{eq:taylor_series_1d}) for the true solution into Equation \ref{eq:Rhie_Chow_1D} results in
\begin{eqnarray}
	\mathcal{D}_{\text {Rhie-Chow}}
	&=& 	\frac{\alpha \bar{K}}{4 \Delta x}	\left[- \bb{u}_{WW} + 4\bb{u}_W - 6\bb{u}_P + 4\bb{u}_{E} - \bb{u}_{EE} \right] \notag \\
	&=& 	\frac{\alpha \bar{K}}{4 \Delta x}
	\Bigg[
	- \bb{U}_P + 2\Delta x \left( \bb{\nabla} \bb{U} \right)_P - 4\frac{\Delta x^2}{2!} \left( \bb{\nabla}^2 \bb{U} \right)_P 
		+ 8\frac{\Delta x^3}{3!} \left( \bb{\nabla}^3 \bb{U} \right)_P \notag \\
	&&\quad\quad + 4\bb{U}_P - 4\Delta x \left( \bb{\nabla} \bb{U} \right)_P + 4\frac{\Delta x^2}{2!} \left( \bb{\nabla}^2 \bb{U} \right)_P
		- 4\frac{\Delta x^3}{3!} \left( \bb{\nabla}^3 \bb{U} \right)_P \notag \\
	&&\quad\quad - 6\bb{U}_P \notag \\
	&&\quad\quad + 4\bb{U}_P + 4\Delta x \left( \bb{\nabla} \bb{U} \right)_P + 4\frac{\Delta x^2}{2!} \left( \bb{\nabla}^2 \bb{U} \right)_P
		+ 4 \frac{\Delta x^3}{3!} \left( \bb{\nabla}^3 \bb{U} \right)_P \notag \\
	&&\quad\quad - \bb{U}_P - 2\Delta x \left( \bb{\nabla} \bb{U} \right)_P - 4\frac{\Delta x^2}{2!} \left( \bb{\nabla}^2 \bb{U} \right)_P 
		+ 8\frac{\Delta x^3}{3!} \left( \bb{\nabla}^3 \bb{U} \right)_P + O(\Delta x^4) \Bigg] \notag \\
	&=& 	\frac{\alpha \bar{K}}{4 \Delta x}	\left[ 16\frac{\Delta x^3}{3!} \left( \bb{\nabla}^3 \bb{U} \right)_P  + O(\Delta x^4) \right] \notag \\
	&=& 	\frac{2}{3} \alpha \bar{K}  \Delta x^2 \left( \bb{\nabla}^3 \bb{U} \right)_P  + O(\Delta x^3)
\end{eqnarray}

\hl{The leading truncation error is proportional to $\Delta x^2$!}
\hl{Did I make a mistake...?}
